### 2020-06-07 report/sections/server/react/efsm.tex
Report: complete chapter 5 EFSM

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
type Props = { factory: SendComponentFactoryFactory };
abstract class S40<State> extends React.Component<Props, State> {
	protected ADD: SendComponentFactory<[number, number]>;
	protected QUIT: SendComponentFactory<[string]>;

@@

We generate an abstract method for each permitted receive
transition, which is used as the receive handler.
This allows the developer to intercept the
message and perform operations -- such as updating
application state -- before the successor state is rendered.
Specifying receive handlers as abstract methods mandate
the developer to provide a concrete implementation.

\begin{lstlisting}[language=javascript,tabsize=2]
enum Labels {
	THANKS = 'THANKS', TERMINATE = 'TERMINATE',
};

type THANKSMessage = {
    label: Labels.THANKS,
    payload: [string],
};
type TERMINATEMessage = {
    label: Labels.TERMINATE,
    payload: [],
};

type Message = | THANKSMessage | TERMINATEMessage;
\end{lstlisting}

\begin{lstlisting}[language=javascript,tabsize=2]
type Props = {
	register: (role: Roles.Peers, handle: ReceiveHandler) => void
};
abstract class S42<S> extends React.Component<Props, S> {
	abstract THANKS(payload1: string): void;
	abstract TERMINATE(): void;

@@

	handle(message: any): State {
		const parsedMessage = JSON.parse(message) as Message;
		switch (parsedMessage.label) {
			case Labels.THANKS:
				return this.THANKS(...parsedMessage.payload);			
			case Labels.TERMINATE:
				return this.TERMINATE(...parsedMessage.payload);
		}
	}
}
\end{lstlisting}

\begin{itemize}
\item receive state is `shown' when waiting for a receive event
\item base class has abstract methods -- one for each possible receive -- so the user must implement these when they extend the abstract base class
\item each abstract method is named after the message label expected to be received, and parameterised by the payload
\item no channel actions are passed to the state -- instead, once the receive state is mounted on the DOM (i.e. rendered), the abstract base class will register the event handlers to the runtime (using the register method passed as props)
\item when the message arrives, the runtime will invoke the registered function (+ forward reference the next section with the edge case that the message arrives before handler registration occurs)
\end{itemize}

\subsection{Terminal States}
\begin{itemize}
\item by convention and design choice, the client gracefully closes the connection with the server role
\end{itemize}

\begin{lstlisting}[language=javascript,tabsize=2]
type Props = { terminate: () => void };
abstract class S41<S> extends React.Component<Props, S> {
	componentDidMount() { this.props.terminate(); }	
}
\end{lstlisting}

>>>>

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2,title=S40.tsx]
type Props = { factory: SendComponentFactoryFactory };
abstract class S40<State> extends React.Component<Props, State> {
	protected ADD: SendComponentFactory<[number, number]>;
	protected QUIT: SendComponentFactory<[string]>;

@@

We generate an \textit{abstract method} 
for each permitted receive
transition, which is used as the receive handler.
This allows the developer to intercept the
message and perform operations -- such as updating
application state -- before the successor state is rendered.
Specifying receive handlers as abstract methods mandate
the developer to provide a concrete implementation.

We demonstrate below how the developer could implement
receive state S42 -- here, the developer manages
stateful application data using the
\textit{sessionStorage} API \cite{SessionStorage}
available on the browser.

\begin{lstlisting}[language=javascript]
class WaitScreen extends S42 {
	render() { return <h1>Waiting</h1>; }
	RES(n: number) {
		sessionStorage.setItem('result', n);
	}
}
\end{lstlisting}

Receive states in \fancyname{ReactMPST}
work in a similar way as with \fancyname{NodeMPST}.
The runtime will invoke the \tmsg{RES} method
implemented by the developer upon receiving the message
through the WebSocket.
In order to do so, the S42 component 
needs to \textit{register}
the handler with the runtime.

Recall that React supports an uni-directional data flow,
so there is no direct way to pass the handler
back up to the parent component.
We overcome this limitation by allowing the
runtime to pass a \texttt{register} function to the
receive state component as a prop; once the receive state
component is mounted on the DOM, it proactively 
registers the handlers with the runtime.

This raises questions about the type signature for the
\texttt{register} function, as the receive handler
identifiers and type signatures differ between receive states.
We adopt the same strategy as with \fancyname{NodeMPST}:
the receive handler registered by the receive state
to the runtime also performs the message parsing,
so all receive states register a ``general handle function'' typed

\begin{lstlisting}[language=javascript,numbers=none]
type MessageHandler = (message: any) => void;
\end{lstlisting}

and define their own \texttt{MessageHandler} to handle
state-specific message parsing and processing.
We construct similar types for the state-specific
labels and message structures.
This is illustrated in \cref{lst:reactreceivetypes}) --
we use state S43 from \cref{fig:adderclientfsm} hereafter
to show how our approach scales with multiple branches.

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
enum Labels { THANKS = 'THANKS', TERMINATE = 'TERMINATE' };
interface THANKSMessage {
    label: Labels.THANKS,
    payload: [string],
};
interface TERMINATEMessage {
    label: Labels.TERMINATE,
    payload: [],
};
type Message = | THANKSMessage | TERMINATEMessage;
\end{lstlisting}
\captionof{lstlisting}{Generated Label and Message Types
in \fancyname{ReactMPST}}
\label{lst:reactreceivetypes}
\end{figure}

We generate the \texttt{handle()} method for each
abstract receive state component for state-specific message
parsing and processing. We show the full generated code
for state S42 in \cref{lst:reactreceivestate}.
Unlike \fancyname{NodeMPST},
the \texttt{handle()} method (\cref{line:reactreceivehandle})
returns the successor state identifier to the runtime
to make the transition.
We leverage the discriminated union from our
generated message types (\cref{lst:reactreceivetypes})
to invoke the correct receive handler (\cref{line:reactreceivediscriminate}).

Again, due to language limitations, we are unable to 
simplify the message processing step as

\begin{lstlisting}[language=javascript,numbers=none]
this[parsedMessage.label](...parsedMessage.payload);
\end{lstlisting}

because this is implicitly expressing a type dependency
over union types,
and \texttt{parsedMessage.label} is unknown at compile-time.

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
type Props = {
	register: (handle: ReceiveHandler) => void
};
abstract class S42<S> extends React.Component<Props, S> {
	abstract THANKS(payload1: string): void;
	abstract TERMINATE(): void;

@@

	handle(message: any): State { (*@\label{line:reactreceivehandle}@*)
		const parsedMessage = JSON.parse(message) as Message;
		switch (parsedMessage.label) { (*@\label{line:reactreceivediscriminate}@*)
			case Labels.THANKS:
				this.THANKS(...parsedMessage.payload);
				return TerminalState.S41;
			case Labels.TERMINATE:
				this.TERMINATE(...parsedMessage.payload);
				return TerminalState.S41;
		}
	}
}
\end{lstlisting}
\captionof{lstlisting}{Generated Code for \trole{Client} 
Receive State S42 in \tprotocol{Adder} protocol}
\label{lst:reactreceivestate}
\end{figure}

\subsection{Terminal States}
To be consistent with the conventions defined for \fancyname{NodeMPST},
the browser endpoint proactively closes the connection at its terminal state.
We generate an abstract React component for terminal states 
to allow the developer to customise what to
\texttt{render} upon protocol termination.

\begin{lstlisting}[language=javascript]
type Props = { terminate: () => void };
abstract class S41<State> extends React.Component<Props, State> {
	componentDidMount() { this.props.terminate(); }	
}
\end{lstlisting}

When the component is mounted on the DOM, it notifies the runtime
(via the received \texttt{terminate} prop) to terminate the session.
